### 2017-03-24 main.tex
Final tweaks
Some minor adjustments based on a final readthrough. It looks great, I think it is ready to go. These changes just reflect suggestions.

* Refer to “AFP” as the “the AFP” (debatable)
* Consistent formatting of times (space before ms)
* Switched one instance of target times to target syllable, as it seemed a bit confusing on my read through
* Rearranged sentence on normalization
* Named Apple’s operating system
* Adjusted some performance information for the MATLAB detector
* Change “timeshifted” to “time-shifted”

Feel free to any of the suggestions or discard them all.

If precise moments in song can reliably be detected quickly enough to
trigger other apparatus during singing, then this consistency of
behaviour allows a variety of experiments.  A common area of study
with song-triggered experiments is the anterior forebrain pathway
(AFP), a homologue of mammalian basal ganglia consisting of a few
distinct brain areas concerned with the learning and production of
song.  For example, stimulation of the lateral magnocellular nucleus
of the anterior nidopallium (LMAN)---the output nucleus of the
AFP---at precisely timed moments during song showed that this area
controls specific variables in song output \cite{Kao2005}.
Song-synchronised stimulation of LMAN and the high vocal centre (HVC)
in one hemisphere or the other showed that control of song rapidly
switches between hemispheres \cite{Wang2008interhemispheric}.
Feedback experiments have shown that Field L and the caudolateral
mesopallium may hold a representation of song against which auditory
signals are compared \cite{Keller2009}.  The disruption by white noise
of renditions of a syllable that were slightly above (or below) the
syllable's average pitch showed that the apparently random natural
variability in songbird motor output is used to drive change in the
song \cite{Tumer2007pitchshift}, and AFP produces a
corrective signal to bias song away from those disruptions
\cite{Andalman2009}. The song change is isolated to within roughly 10
milliseconds (ms) of the stimulus, and the shape of the learned
response can be predicted by a simple mechanism
\cite{Charlesworth2011learning}. AFP transfers the error signal to the
robust nucleus of the arcopallium (RA) using NMDA-receptor--mediated
glutamatergic transmission \cite{Warren2011}. The course of song
recovery after applying such a pitch-shift paradigm showed that the
caudal medial nidopallium is implicated in memorising or recalling a
recent song target, but in neither auditory processing nor directed
motor learning \cite{Canopoli2014}.

@@

A variety of syllable-triggering systems have been used, but few have
been documented or characterised in detail.  In 1999,
detection was achieved by a group of IIR filters with hand-tuned logical
operators \cite{Leonardo1999}.  The system had a latency of 50 or 100 ms,
and accuracy and jitter were not reported.  As access to
computational resources has improved, approaches have changed: in
2009, hand-tuned filters were implemented on a Tucker-Davis Technologies digital signal processor, bringing latency down to around 4 ms \cite{Andalman2009}. But as with other filter-bank techniques, it
is not strictly a syllable detector but rather a pitch and timbre
detector---it cannot identify a frequency sweep, or distinguish a
short chirp from a long one---and thus requires careful selection of
target syllables.  Furthermore, the method is neither inexpensive nor,
based on our experience with a similar technique, accurate.
2009 saw the application of a neural network to a spectral image of
song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8-ms
segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
false-positive rates of up to 2\% and 4\% respectively.  Hardware
requirements and ease of use were not reported.  In 2013, spectral
images of template syllables were compared to song using a correlation coefficient \cite{Skocik2013}.  With a
fast desktop (Intel i7 six-core) running Linux and equipped with a
National Instruments data acquisition card, it boasts a hardware-only
(not accounting for the time taken to compute a match with a syllable)
latency and jitter of just a few microseconds, and the detection
computation they use should not much increase that.  They reported
false-negative rates around 4\% and 7\% for zebra finches and
Bengalese finches respectively, measured on a small dataset.  In much
other work, a syllable detector is alluded to, but not described.

@@

Training targets $y_t$ are vectors with one element for each desired
detection syllable.  That element is, roughly, 1 if the input vector
comes from the target time ($t=t^*$), 0 otherwise, for each target
syllable (of which there may be any number, although they increase
training time, and in our implementations the number of distinct
output pulses is constrained by hardware).  Since the song alignment
may not be perfect, and due to sample aliasing, a
strict binary target may ask the network to learn that, of two
practically identical frames, one should be a match and the other
not. Thus it is preferable to spread the target in time, such that at
the target moment it is 1, and at neighbouring moments it is
nonzero. We found that a Gaussian smoothing kernel around the target
time with a standard deviation of 2 ms serves well.

@@

In order to present consistent and meaningful inputs to the neural
network, we normalise the incoming data stream so that changes in the
song structure of the sound are emphasised over changes in volume, and
to maximise the effectiveness of the neural network's training
algorithm.

@@

This detector uses the Swift programming language and Core Audio
interface included in Apple's operating systems.

@@

The most versatile implementation, MATLAB runs on a variety of
hardware and operating systems, and is perhaps the easiest to modify.
While it did not perform as well on our test system as the other
implementations, the convenience may outweigh the timing
performance penalty for some experiments.  Key to minimising jitter is
the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
the smallest buffer size that did not result in read overruns was
about 4 ms.

As with the Swift detector, it is also possible to modify the
MATLAB implementation to generate the TTL pulse from a microcontroller
(Arduino) controlled via a serial over USB interface. This eliminates
the double buffer required when sending triggers via the audio
interface, reducing latency by half.

@@

We measure accuracy and characterise timing for 6 randomly chosen
points in one bird's song (lny64), and present accuracy results for
seven more birds (the detector's timing does not appreciably vary among
birds). The results presented here are representative of real-world
performance in feedback experiments on many birds in our lab.

@@

We define the accuracy of the network based on its classification
performance per frame. In order to avoid the apparent problem of every
non-detected non-matching frame counting as a true negative, we also
tried defining accuracy on a per-song basis, such that a song-length
sample without the target syllable counted as a single true negative.
Computing the optimal output thresholds on a per-frame rather than a
per-song basis resulted in higher thresholds and thus a slight
reduction in both the false-positive and true-positive rates
(reversible by increasing $C_n$), while also providing a valid
definition of false-positive rate for data streams that had not been
segmented into song-sized chunks and thus allowing easy interpretation
of false-positive rates from calls, cage noise, etc.

@@

An example of the detector's output for three of these syllables is
shown in \fig{fig:detection_raster}.  The total audio energy of each
song is shown as a single grayscale row.  By sorting according to time
of each syllable's detection, the rest of the song is shown in the
context of the timing for that moment.  This gives an intuition of the
timing variability within each song and across songs (which is
responsible for a small amount of our measured jitter, but note that few of the songs are significantly timeshifted).

>>>>

If precise moments in song can reliably be detected quickly enough to
trigger other apparatus during singing, then this consistency of
behaviour allows a variety of experiments.  A common area of study
with song-triggered experiments is the anterior forebrain pathway
(AFP), a homologue of mammalian basal ganglia consisting of a few
distinct brain areas concerned with the learning and production of
song.  For example, stimulation of the lateral magnocellular nucleus
of the anterior nidopallium (LMAN)---the output nucleus of the
AFP---at precisely timed moments during song showed that this area
controls specific variables in song output \cite{Kao2005}.
Song-synchronised stimulation of LMAN and the high vocal centre (HVC)
in one hemisphere or the other showed that control of song rapidly
switches between hemispheres \cite{Wang2008interhemispheric}.
Feedback experiments have shown that Field L and the caudolateral
mesopallium may hold a representation of song against which auditory
signals are compared \cite{Keller2009}.  The disruption by white noise
of renditions of a syllable that were slightly above (or below) the
syllable's average pitch showed that the apparently random natural
variability in songbird motor output is used to drive change in the
song \cite{Tumer2007pitchshift}, and the AFP produces a
corrective signal to bias song away from those disruptions
\cite{Andalman2009}. The song change is isolated to within roughly 10
milliseconds (ms) of the stimulus, and the shape of the learned
response can be predicted by a simple mechanism
\cite{Charlesworth2011learning}. The AFP transfers the error signal to the
robust nucleus of the arcopallium (RA) using NMDA-receptor--mediated
glutamatergic transmission \cite{Warren2011}. The course of song
recovery after applying such a pitch-shift paradigm showed that the
caudal medial nidopallium is implicated in memorising or recalling a
recent song target, but in neither auditory processing nor directed
motor learning \cite{Canopoli2014}.

@@

A variety of syllable-triggering systems have been used, but few have
been documented or characterised in detail.  In 1999,
detection was achieved by a group of IIR filters with hand-tuned logical
operators \cite{Leonardo1999}.  The system had a latency of 50 or 100 ms,
and accuracy and jitter were not reported.  As access to
computational resources has improved, approaches have changed: in
2009, hand-tuned filters were implemented on a Tucker-Davis Technologies digital signal processor, bringing latency down to around 4 ms \cite{Andalman2009}. But as with other filter-bank techniques, it
is not strictly a syllable detector but rather a pitch and timbre
detector---it cannot identify a frequency sweep, or distinguish a
short chirp from a long one---and thus requires careful selection of
target syllables.  Furthermore, the method is neither inexpensive nor,
based on our experience with a similar technique, accurate.
2009 saw the application of a neural network to a spectral image of
song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8 ms
segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
false-positive rates of up to 2\% and 4\% respectively.  Hardware
requirements and ease of use were not reported.  In 2013, spectral
images of template syllables were compared to song using a correlation coefficient \cite{Skocik2013}.  With a
fast desktop (Intel i7 six-core) running Linux and equipped with a
National Instruments data acquisition card, it boasts a hardware-only
(not accounting for the time taken to compute a match with a syllable)
latency and jitter of just a few microseconds, and the detection
computation they use should not much increase that.  They reported
false-negative rates around 4\% and 7\% for zebra finches and
Bengalese finches respectively, measured on a small dataset.  In much
other work, a syllable detector is alluded to, but not described.

@@

Training targets $y_t$ are vectors with one element for each desired
detection syllable.  That element is, roughly, 1 if the input vector
matches the target syllable ($t=t^*$), 0 otherwise, for each target
syllable (of which there may be any number, although they increase
training time, and in our implementations the number of distinct
output pulses is constrained by hardware).  Since the song alignment
may not be perfect, and due to sample aliasing, a
strict binary target may ask the network to learn that, of two
practically identical frames, one should be a match and the other
not. Thus it is preferable to spread the target in time, such that at
the target moment it is 1, and at neighbouring moments it is
nonzero. We found that a Gaussian smoothing kernel around the target
time with a standard deviation of 2 ms serves well.

@@

In order to present consistent and meaningful inputs to the neural
network and to maximise the effectiveness of the neural network's training
algorithm, we normalise the incoming data stream so that changes in the
song structure of the sound are emphasised over changes in volume.

@@

This detector uses the Swift programming language and Core Audio
interface included in Apple's Mac OS X operating systems.

@@

The most versatile implementation, MATLAB runs on a variety of
hardware and operating systems, and is perhaps the easiest to modify.
While it did not perform as well as the other
implementations, the convenience may outweigh the timing
performance penalty for some experiments.  Key to minimising jitter is
the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
the smallest buffer size that did not result in read overruns was
about 4 ms.

As with the Swift detector, it is also possible to modify the
MATLAB implementation to generate the TTL pulse from a microcontroller
(Arduino) controlled via a serial over USB interface. This eliminates
the double buffer required when sending triggers via the audio
interface, reducing latency by close to half.

@@

We measure accuracy and characterise timing for 6 randomly chosen
points in one bird's song (lny64), and present accuracy results for
seven more birds (the detector's timing does not appreciably vary across
birds). The results presented here are representative of real-world
performance in feedback experiments on many birds in our lab.

@@

We define the accuracy of the network based on its classification
performance per frame. In order to avoid the apparent problem of every
non-detected non-matching frame counting as a true negative, we also
tried defining accuracy on a per-song basis, such that a song-length
sample without the target syllable counted as a single true negative.
Computing the optimal output thresholds on a per-frame rather than a
per-song basis resulted in higher thresholds and thus a slight
reduction in both the false-positive and true-positive rates
(reversible by increasing $C_n$), while also providing a valid
definition of the false-positive rate for data streams that had not been
segmented into song-sized chunks and thus allowing easy interpretation
of false-positive rates from calls, cage noise, etc.

@@

An example of the detector's output for three of these syllables is
shown in \fig{fig:detection_raster}.  The total audio energy of each
song is shown as a single grayscale row.  By sorting according to time
of each syllable's detection, the rest of the song is shown in the
context of the timing for that moment.  This gives an intuition of the
timing variability within each song and across songs (which is
responsible for a small amount of our measured jitter, but note that few of the songs are significantly time-shifted).
